Which Rules Entail this Fact? \newline An Efficient Approach Using RDBMSs

### 2022-12-17 Paper/Overleaf/paper.tex
Paper - Final Version #2

\begin{document}
%%% Mehrere Autoren werden durch \and voneinander getrennt.
%%% Die Fußnote enthält die Adresse sowie eine E-Mail-Adresse.
%%% Das optionale Argument (sofern angegeben) wird für die Kopfzeile verwendet.
\title[Which Rules Entail this Fact? An Efficient Approach Using RDBMSs]{Which Rules Entail this Fact? An Efficient Approach Using RDBMSs}
%%% \subtitle{From Facts to Rules using Relational Databases} % if needed
\author[Tim Gutberlet \and Janik Sauerbier]
{Tim Gutberlet\footnote{University of Mannheim, \email{tim.gutberlet@students.uni-mannheim.de}} \and
Janik Sauerbier\footnote{University of Mannheim, \email{janik.sauerbier@students.uni-mannheim.de}}}
\startpage{1} % Beginn der Seitenzählung für diesen Beitrag / Start page
\editor{B. König-Ries et al.} % Names of Editors
\booktitle{Datenbanksysteme für Business, Technologie und Web (BTW 2023)} % Name of book title
\yearofpublication{2023}
%%%\lnidoi{18.18420/provided-by-editor-02} % if known
\maketitle

@@

The given KG is stored in one table of the relational database (\textit{kg\_table} with columns “sub” for subjects, “rel” for relations and “obj” for objects). The basic idea behind our approach is creating SQL queries which check whether a certain rule entails a certain target fact. Those individual queries are then combined using \textit{UNION ALL} operations over all rules, where the head can unify with the target fact. To improve this basic idea, we employ several optimizations listed below. This would be the query for rule (\ref{eq:rule_2}) and the target fact given in the preliminaries:

@@

\subsection{Advanced Rule Pre-filtering} 
Firstly, we only consider rules that can unify with the target fact for the query. This can be done by naively testing each rule. However, it can also be done more efficiently. The easiest way is storing each rule under its head as key in a hash map. The head stays unchanged, but we use “X” and “Y” as variable descriptors.

Let (s, r, o) be a target triple with the entities s and o and the relation r. Then we only use the rules stored under the keys \textit{(X, r, Y)}, \textit{(s, r, X)}, \textit{(X, r, o)} or \textit{(s, r, o)} for the query. Additionally, when it is true that s=o for the target, we also use the key \textit{(X, r, X)}. Instead of looping through n rules naively in O(n), this allows for initial filtering of the rules in O(1).

@@

The top x\% of rules are then pre-computed, by calculating all potential assignments for variables in the rule head that form a grounding together with a set of facts from the KG. We then store these combinations in a table for the rule. If the subject or the object is a variable and the other one a constant, then the table only contains the column for the variable. For the implementation of the tables, we use materialized views to \(SELECT\) from the KG with all the conditions we already know from the rule. Consequently, the \(CREATE\) statement for pre-computing rule (\ref{eq:rule_2}) and the query for the target fact from the preliminaries would be:

@@

\begin{table}[t]
\centering
\begin{tabular}{l|lll|lll}
\toprule
 & 10s & 50s & 100s & 1 body atom & 2 body atoms & 3 body atoms\\
\midrule
YAGO3-10 & 27k & 107k & 165k & 82.8\% & 11.2\% & 6.0\%\\
WN18RR & 6k & 20k & 30k & 33.0\% & 23.9\% & 43.1\%\\
FB15k-237 & 33k & 135k & 252k & 53.9\% & 26.2\% & 19.9\%\\
\bottomrule
\end{tabular}
\caption{Composition of the rules (\# of rules | avg. share by \# of body atoms)}
\label{tab:rules}
\end{table}

>>>>

\begin{document}
%%% Mehrere Autoren werden durch \and voneinander getrennt.
%%% Die Fußnote enthält die Adresse sowie eine E-Mail-Adresse.
%%% Das optionale Argument (sofern angegeben) wird für die Kopfzeile verwendet.
\title[Which Rules Entail this Fact? An Efficient Approach Using RDBMSs]{Which Rules Entail this Fact? \newline An Efficient Approach Using RDBMSs}
%%% \subtitle{From Facts to Rules using Relational Databases} % if needed
\author[Tim Gutberlet \and Janik Sauerbier]
{Tim Gutberlet\footnote{University of Mannheim, \email{tim.gutberlet@students.uni-mannheim.de}} \and
Janik Sauerbier\footnote{University of Mannheim, \email{janik.sauerbier@students.uni-mannheim.de}}}
\startpage{1} % Beginn der Seitenzählung für diesen Beitrag / Start page
\editor{B. König-Ries et al.} % Names of Editors
\booktitle{Datenbanksysteme für Business, Technologie und Web (BTW 2023)} % Name of book title
\yearofpublication{2023}
%%%\lnidoi{18.18420/provided-by-editor-02} % if known
\maketitle

@@

The given KG is stored in one table of the relational database (\textit{kg\_table} with columns \textit{sub} for subjects, \textit{rel} for relations and \textit{obj} for objects). The basic idea behind our approach is creating SQL queries which check whether a certain rule entails a certain target fact. Those individual queries are then combined using \textit{UNION ALL} operations over all rules, where the head can unify with the target fact. To improve this basic idea, we employ several optimizations listed below. This would be the query for rule (\ref{eq:rule_2}) and the target fact given in the preliminaries:

@@

\subsection{Advanced Rule Pre-filtering} 
Firstly, we only consider rules that can unify with the target fact for the query. This can be done by naively testing each rule. However, it can also be done more efficiently. The easiest way is storing each rule under its head as key in a hash map. The head stays unchanged, but we use “\textit{X}” and “\textit{Y}” as variable descriptors.

Let \textit{(s, r, o)} be a target triple with the entities \textit{s} and \textit{o} and the relation \textit{r}. Then we only use the rules stored under the keys \textit{(X, r, Y)}, \textit{(s, r, X)}, \textit{(X, r, o)} or \textit{(s, r, o)} for the query of the target triple. Additionally, when it is true that \(s=o\), we also use the key \textit{(X, r, X)}. Instead of looping through n rules naively in O(n), this allows for pre-filtering of the rules in O(1).

@@

The top x\% of rules are then pre-computed, by calculating all potential assignments for variables in the rule head that form a grounding together with a set of facts from the KG. We then store these combinations in a table for the rule. If the subject or the object is a variable and the other one a constant, then the table only contains the column for the variable. For the implementation, we use indexed materialized views to \(SELECT\) from the KG with all the conditions we already know from the rule. Consequently, the \(CREATE\) statement for pre-computing rule (\ref{eq:rule_2}) and the query for the target fact from the preliminaries would be:

@@

\begin{table}[t]
\centering
\begin{tabular}{l|lll|lll}
\toprule
 & 10s & 50s & 100s & 1 body atom & 2 body atoms & 3 body atoms\\
\midrule
YAGO3-10 & 27k & 107k & 165k & 82.8\% & 11.2\% & 6.0\%\\
WN18RR & 6k & 20k & 30k & 33.0\% & 23.9\% & 43.1\%\\
FB15k-237 & 33k & 135k & 252k & 53.9\% & 26.2\% & 19.9\%\\
\bottomrule
\end{tabular}
\caption{Composition of the rules (\# of rules | avg. share of rules by \# of body atoms)}
\label{tab:rules}
\end{table}
